---
title: "Most Harmful and Economic Weather Events Between 1950 and 2011 in USA"
author: "Deyban Pérez"
date: "February 22, 2018"
output: 
  pdf_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
rm(list = ls())
source("functions.r")
InstallLoadPackages("tidyr")
InstallLoadPackages("dplyr")
InstallLoadPackages("lubridate")

```


## Data Processing

The [data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2) used is extracted from U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database who has the mainly weather events ocurred from 1950 to 2011.

```{r load_data, cache=TRUE}

if(!file.exists("./Data/dataset.bz2"))
{
  data_url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
  dir.create(path = "./Data")
  data = download.file(data_url, "./Data/dataset.bz2", method = "libcurl")
}

dataset = read.csv("./Data/dataset.bz2", sep = ",")
```

Takes a long of time to the dataset be loaded, that is the reasen why this chunk of code is marked as cache =  TRUE. Acoordying with the [documentation](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf) especifically pages 5-15 the columns of interest related to the assigment that envolves: harmful, economic expenses and event type, we just need the following columns from the dataset:

1. STATE__: int, state ID
2. BGN_DATE: date, beginning date of the event
3. FATALITIES: int, number of people who dead
4. INJURIES: int, number of people who suffered an injury
5. PROPDMG: double, property damages quantity
6. PROPDMGEXP: char, K (thousand), M (million), B (billion)
7. CROPDMG: double, crop damages quantity
8. CROPDMGEXP: char, K (thousand), M (million), B (billion)

```{r}
dataset = as.tbl(dataset)
dataset = select(dataset, STATE__:BGN_DATE, FATALITIES:CROPDMGEXP)
dataset$BGN_DATE = parse_date_time(dataset$BGN_DATE, orders = c("mdyHMS"))
dataset$STATE__ = as.factor(dataset$STATE__)
dataset = arrange(dataset, BGN_DATE)

glimpse(dataset)
str(dataset)
dim(dataset)
names(dataset)
```

## Results

